Stochastic Variance Reduced Policy Gradient

### 2018-02-02 svrpg.tex
completed introduction

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator*{\Cov}{\mathbb{C}ov}
\DeclareMathOperator*{\Var}{\mathbb{V}ar}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
\newcommand{\dotprod}[2]{\left\langle#1,#2\right\rangle}
\newcommand{\dnabla}{\nabla\!\!\!\!\nabla}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla J(#1)}
\newcommand{\gradApp}[2]{\widehat{\nabla}_{#2}J(#1)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%Specific
\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\mid s}\left[#2\right]}
\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\mid s}\left(#2,#3\right)}
\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\mid s}\left[#2\right]}
\newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
\newcommand{\gradIdeal}[1]{\dnabla J(#1)}
\newcommand{\VARRF}{V}
\newcommand{\GRADLOG}{G}
\newcommand{\VARIS}{W}
\newcommand{\HESSLOG}{F}
%%%%%%

@@

\section{Introduction}
%Similarities between RL and SL
On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a given task. The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal. This makes RL fundamentally different from Supervised Learning (SL), even in its online version, where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels). However, the two approaches share many challenges and, as a consequence, solution concepts. The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behaviour directly. Moreover, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.

% SL needs SVRG so RL
In a typical discriminative supervised learning setting, an objective function $J(x\mid\vtheta)$ has to be optimized \wrt to a parameter vector $\vtheta$. The training set $\mathcal{D}_{train}$ is the set of data samples $x$ for which the objective (and its gradient) can be evaluated. In most practical cases, $\mathcal{D}_{train}$ is only a subset of the set $\mathcal{D}$ of all the $x$ of interest, hence we resort to stochastic optimization of the expected value of $J$ over an unknown data distribution. The special case in which $\mathcal{D}_{train} = \mathcal{D}$ is known as Empirical Risk Minimization (ERM), and allows for deterministic optimization. Even in this case, stochastic optimization is often used for reasons of efficiency. The idea is to evaluate the objective on a random subset of the available data to obtain an approximate gradient direction. This can be much less expensive than taking into account all the data for all parameter updates. However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that speed and variance must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG, SVRG and SAGA offer better ways of achieving the optimal trade-off between variance and speed, both in theory and practice. Although designed explicitly for ERM, this algorithms address a problem that is equally important, if not more critical, in more general machine learning problems.
In Reinforcement Learning, as in online learning n general, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment. In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative, or a complement, to traditional methods such as batch learning.

\todomat{work in progress}
% Potential advantages of SVRG
% Difficulties in RL \wrt SL

% Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.

\section{Preliminaries}
In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

@@

\subsection{Stochastic Variance-Reduced Gradient}
Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\cdot|\vtheta)$:
\begin{align*}
        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta)\right\}.
\end{align*}
This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $g_i(\vtheta) = g(x_i|\vtheta)$). 
%% In this case, we adopt the following, more meaningful notation:
%% \begin{align*}
%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \mid \vtheta).
%% \end{align*}
A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\todopirout{Is it $\ell_2$-norm?} 
This can be too expensive for large values of $N$. Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$. 
\todopir{What is the reference?}
Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito}, 
The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.

@@

\begin{algorithm}[tb]
	\caption{SVRG}
	\label{alg:svrg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
		\FOR{$t=0$ {\bfseries to} $m-1$}
		\STATE Pick $i$ uniformly at random from $[1,N]$
		\STATE \begin{align*}
			\blacktriangledown f(\vtheta_t^{s+1}) = 
			&\nabla f(\tilde{\vtheta}^s) \\ + 
			&\nabla g(x_i\mid\vtheta_t^{s+1}) -
			\nabla g(x_i \mid \tilde{\vtheta}^{s+1})
		\end{align*}
		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown f(\vtheta_t^{s+1})$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries Convex case: return} $\vtheta_{m-1}^S$
		\STATE {\bfseries Non-Convex case: return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
	\end{algorithmic}
\end{algorithm}

\section{Algorithm}\label{sec:algo}

@@

	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),

@@

where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$. 

@@

\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\

@@

\nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i)\right].

@@

\begin{algorithm}[tb]
	\caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$, mini-batch size B
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\mid\tilde{\vtheta}^{s})$
		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\mid\tilde{\vtheta}^{s}}\Reward(\tau_j)$
		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
		\FOR{$t=0$ {\bfseries to} $m-1$}
		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\mid\vtheta_t^{s+1})$
		\STATE 
		\begin{align*}
		\blacktriangledown J(\vtheta_t^{s+1}) = 
		&\gradApp{\tilde{\vtheta}^s}{N} \\
		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
		\score{}{\tau_i\mid\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
		&\left. - \omega(\tau_i)\score{}{\tau_i \mid \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
		\end{align*}
		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown J(\vtheta_t^{s+1})$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
	\end{algorithmic}
\end{algorithm}

@@

\section{Hyperparameters}\label{sec:stopping}
In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot. The first one is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots. The second one is essetial to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.

A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter. Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update. The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates. By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations. These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation. For the above mentioned reasons let us denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.

We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too. Therefore, if $m$ is not properly set the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.

We have seen, sperimentally, that a fixed epoch size is not the optimal choice, but an adaptive one would better exploit the total amount of data at hand. Hence, we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot. This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
More precisely we stop performing sub-iterations when:

@@

These two weights ${\bf{\it N}}$ and ${\bf{\it B}}$ represent, respectively, the number of trajectories used for a snapshot and the number of trajectories used for an update of the parameters in a sub-iteration. Therefore, in this stopping condition, we use them to take into account the trajectories efficiency. The less trajectories we use to perform an update the better it is from a data usage point of view. The variance control offered by this adaptive stopping condition lies within the value of the two different learning rate proposed by adam. Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance. Hence it could be better to perform a new snapshot so to mitigate this issue.

\section{Practical Variants}

We introduce a term with importance weights to guarantee proper sampling of trajectories.
This term's variance is subjected to the variance of importance weights. 
There are some methods to reduce this variance.

As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance. The off-policy gradient estimator becomes:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
\end{align*} 
where:
\begin{align*}
\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
 }{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
\end{align*}
\begin{align*}
\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\mid s_k^n})
}{\pi_{\vtheta^{'}}({a_k^n\mid s_k^n})}
\end{align*}

\section{Related Work}

@@

	&= \mathbb{E}_{t\mid s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}

@@

	\mathbb{E}_{t\mid s}\left[\left\|

>>>>

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator*{\Cov}{\mathbb{C}ov}
\DeclareMathOperator*{\Var}{\mathbb{V}ar}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
\newcommand{\dotprod}[2]{\left\langle#1,#2\right\rangle}
\newcommand{\dnabla}{\nabla\!\!\!\!\nabla}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \vert s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla J(#1)}
\newcommand{\gradApp}[2]{\widehat{\nabla}_{#2}J(#1)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%Specific
\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\vert s}\left[#2\right]}
\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\vert s}\left(#2,#3\right)}
\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\vert s}\left[#2\right]}
\newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
\newcommand{\gradIdeal}[1]{\dnabla J(#1)}
\newcommand{\VARRF}{V}
\newcommand{\GRADLOG}{G}
\newcommand{\VARIS}{W}
\newcommand{\HESSLOG}{F}
%%%%%%

@@

\section{Introduction}
%Similarities between RL and SL
On a very general level, machine learning addresses the problem of an agent that must select the right actions to solve a task. The approach of Reinforcement Learning~\citep{sutton1998reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal. This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels). However, the two approaches share many challenges and tools. The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behaviour directly. Moreover, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks.
% SL needs SVRG so RL
In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$. The set of data that are available for training is often a subset of all the data of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution. When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM). Even in this case, stochastic optimization is often used for reasons of efficiency. The idea of stochastic gradient descent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction. At the level of the single iteration, this can be much less expensive than taking into account all the data. However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of addressing this trade-off, with significant results both in theory and practice. Although designed explicitly for ERM, this algorithms address a problem that affects more general machine learning problems. 
In RL, stochastic optimization is rarely a matter of choice, since data must be actively sampled by interacting with an initially unknown environment. In this scenario, containing the variance of the estimates is a problem that cannot be avoided, which makes variance-reduced algorithms very interesting.
% Potential advantages of SVRG
Among RL approaches, policy gradient is the one that bears the closest similarity to SL solutions. The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient descent. Compared to other applications of SGD, the cost of collecting samples can be very high since it requires to interact with the environment. This makes SVRG-like methods potentially much more efficient than, \eg batch learning.
% Difficulties in RL \wrt SL
Unfortunately, RL has a series of difficulties that are not present in ERM. First, in SL the objective can often be designed to be strongly convex. This is not possible in RL, so we have to deal with non-convexities. Then, as mentioned, the dataset is not initially available and may even be infinite, which  makes approximations unavoidable. This also rules out SAG and SAGA because of their storage requirements, which leaves SVRG as the most promising choice. Finally, the distribution used to sample data is not under direct control of the algorithm designer, but is a function of policy parameters that changes over time as the policy is optimized, which is a form of nonstationarity.

% Related work

%Paper structure
After providing a background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address the aforementioned difficulties to design SVRPG, a variant of SVRG for the policy gradient framework. In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL. In Section \ref{sec:stopping} we suggests how to set the hyperparameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm. Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL benchmark tasks.

\section{Preliminaries}\label{sec:pre}
In this section we provide the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

@@

\subsection{Stochastic Variance-Reduced Gradient}
Finite-sum optimization is the problem of maximizing an objective function $f(\vtheta)$ which can be decomposed into the sum or average of a finite number of functions $g_i(\cdot|\vtheta)$:
\begin{align*}
        \max_{\vtheta} \left\{ f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g_i(\vtheta)\right\}.
\end{align*}
This kind of optimization is very common in machine learning, where each $g_i$ may correspond to a data sample $x_i$ from a dataset $\mathcal{D}_N$ of size $N$ (\ie $g_i(\vtheta) = g(x_i|\vtheta)$). 
%% In this case, we adopt the following, more meaningful notation:
%% \begin{align*}
%% \max_{\vtheta} f(\vtheta) = \frac{1}{N}\sum_{i=1}^{N}g(x_i \vert \vtheta).
%% \end{align*}
A common requirement is that $g$ must be smooth and {\color{red} concave} in $\vtheta$.\footnote{Note that we are considering a maximization problem instead of the classical minimization one.} Under this hypothesis, full gradient ascent \cite{cauchy1847methode} with a constant step size achieves an Incremental First-order Oracle (IFO) complexity~\citep{agarwal2014lower} of $O(\nicefrac{N}{\epsilon})$, \ie it requires $k = O(\nicefrac{N}{\epsilon})$ gradient computations to achieve $\norm[]{\nabla_{\vtheta}f(\vtheta_k)}^2\leq\epsilon$.\todopirout{Is it $\ell_2$-norm?} 
This can be too expensive for large values of $N$. Stochastic Gradient (SG) ascent~\citep[\eg][]{robbins1951stochastic,bottou2004large} overcomes this problem by sampling a single sample $x_i$ per iteration, but a vanishing step size is required to control the variance introduced by sampling. The resulting IFO complexity is $O(\nicefrac{1}{\epsilon^2})$ in expectation, which does not depend on $N$, but has a worse dependency on $\epsilon$. 
\todopir{What is the reference?}
Starting from SAG, a series of variations to SG have been proposed to achieve a better trade-off between convergence speed and cost per iteration: \eg SAG~\citep{roux2012stochastic}, SVRG~\cite{johnson2013accelerating}, SAGA~\cite{defazio2014saga} and Finito~\cite{defazio2014finito}, 
The common idea is to reuse past gradient computations to reduce the variance of the current estimate.
In particular, Stochastic Variance-Reduced Gradient (SVRG) is often preferred to other similar methods for its limited storage requirements, which can become a problem when deep and/or wide neural networks are employed.

@@

\begin{algorithm}[tb]
	\caption{SVRG}
	\label{alg:svrg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
		\FOR{$t=0$ {\bfseries to} $m-1$}
		\STATE Pick $i$ uniformly at random from $[1,N]$
		\STATE \begin{align*}
			\blacktriangledown f(\vtheta_t^{s+1}) = 
			&\nabla f(\tilde{\vtheta}^s) \\ + 
			&\nabla g(x_i\vert\vtheta_t^{s+1}) -
			\nabla g(x_i \vert \tilde{\vtheta}^{s+1})
		\end{align*}
		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown f(\vtheta_t^{s+1})$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries Convex case: return} $\vtheta_{m-1}^S$
		\STATE {\bfseries Non-Convex case: return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
	\end{algorithmic}
\end{algorithm}

\section{Algorithm}\label{sec:alg}

@@

	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),

@@

where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\vert\tilde{\vtheta})}{\pi(\tau\vert\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$. 

@@

\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\

@@

\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].

@@

\begin{algorithm}[tb]
	\caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m-1}^0$, batch size $N$, mini-batch size B
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE Sample N trajectories $\tau_j$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\vert\tilde{\vtheta}^{s}}\Reward(\tau_j)$
		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m-1}^s$
		\FOR{$t=0$ {\bfseries to} $m-1$}
		\STATE Sample $B$ trajectories $\tau_i$ from 				$\pi(\cdot\vert\vtheta_t^{s+1})$
		\STATE 
		\begin{align*}
		\blacktriangledown J(\vtheta_t^{s+1}) = 
		&\gradApp{\tilde{\vtheta}^s}{N} \\
		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
		\end{align*}
		\STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha\blacktriangledown J(\vtheta_t^{s+1})$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
	\end{algorithmic}
\end{algorithm}

@@

\section{Hyperparameters}\label{sec:stopping}
In Reinforcement learning we have to take into account the gradient's variance.
\todomat{$\uparrow$ the opening of the section could connect better with previous sections. Declare which meta-parameters will be addressed in the section}
A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter. Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
We use two different adam hyperparameters: one for updates tied to snapshot points and the other tied to the updates of sub-iterations points.
\todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules. Use notation from previous sections, \eg snaphot is usually associated with $\tilde{\vtheta}$} 
This is because we are dealing with two different variance magnitudes.
\todomat{$\uparrow$ the reason is this, but try to spend more words on it}
Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.

As shown in Section \ref{sec:conv}, the performance of the SVRPG algorithm is sensitive to the epoch size $m$.
As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
A fixed epoch size is not the optimal choice.
Few sub-iterations lead to few updates, many update lead to increases of variance.

In order to cope with this behaviour we introduce an adaptive stopping condition: we perform a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot. This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
\todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
\todobep{I tried to solve the issue here:hope it's better now}
We stop with inner iterations if:

@@

\todomat{why? should mention how this formula combines trajectory efficiency and variance control}
After that we sample N trajectories for the next Snapshot point.

\section{Practical Variants}\label{sec:prac}
\todomat{Missing opening to section}
As reported in Section \ref{subsec:PolicyGradient} we use the G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance. The off-policy gradient estimator becomes:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
\end{align*} 
where:
\begin{align*}
\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
 }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\begin{align*}
\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
\section{Related Work}

@@

	&= \mathbb{E}_{t\vert s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}

@@

	\mathbb{E}_{t\vert s}\left[\left\|
